---
title: "Spotify 2023 Streaming Analysis"
author: "[ShiriCodes](https://github.com/ShiriCodes/spotify-2023-data-analysis)"
date: "2025-10-26"
description: "Statistical analysis of Spotify's most streamed songs in 2023. Exploring danceability trends, BPM impact on popularity, energy-valence correlations, and mode effects using R statistical methods. Comprehensive data analysis with product insights for music streaming platforms."
output:
  html_document:
    toc: true
    toc_float: true
    theme: lumen
    highlight: tango
    code_folding: hide
---

<meta name="description" content="Statistical analysis of Spotify's 2023 top streaming songs. Data-driven insights on danceability, tempo, energy, valence, and musical mode using R, ggplot2, and statistical testing.">
<meta name="keywords" content="Spotify analysis, music data analysis, streaming statistics, R programming, data visualization, danceability analysis, BPM analysis, audio features, statistical testing, product analytics">
<meta name="author" content="ShiriCodes">
<meta property="og:title" content="Spotify 2023 Streaming Analysis | Data Analysis by ShiriCodes">
<meta property="og:description" content="Statistical analysis revealing trends in Spotify's top 2023 songs: danceability patterns, tempo impact, and mood correlations with product recommendations.">
<meta property="og:type" content="article">

```{css, echo=FALSE}
/* ====== FONTS ====== */
@import url('https://fonts.googleapis.com/css2?family=Inter:wght@400;500;600;700&family=Playfair+Display:wght@600;700&display=swap');

/* ====== PAGE BACKGROUND ====== */
html, body {
  height: 100%;
  margin: 0;
  background: linear-gradient(135deg, #e8f4f8 0%, #f6f9fc 50%, #fef9f3 100%);
  background-attachment: fixed;
}

/* ====== MAIN CONTENT (Glass Card) ====== */
#content {
  font-family: "Inter", "Helvetica Neue", Helvetica, Arial, sans-serif;
  color: #2b2b2b;
  line-height: 1.7;
  font-size: 15px;

  background: rgba(255, 255, 255, 0.65);
  backdrop-filter: blur(16px) saturate(180%);
  -webkit-backdrop-filter: blur(16px) saturate(180%);
  border-radius: 16px;
  border: 1px solid rgba(255, 255, 255, 0.3);
  box-shadow: 0 8px 32px rgba(0, 0, 0, 0.08);

  max-width: 1000px;
  margin: 50px auto;
  padding: 50px 70px;
}

/* ====== HEADINGS ====== */
h1 {
  font-family: "Playfair Display", Georgia, serif;
  color: #1DB954;
  font-weight: 700;
  font-size: 2.5em;
  margin-bottom: 10px;
  text-shadow: 0 2px 4px rgba(0,0,0,0.06);
}

h2 {
  font-family: "Playfair Display", Georgia, serif;
  color: #004b66;
  font-weight: 600;
  font-size: 1.8em;
  border-bottom: 2px solid rgba(29, 185, 84, 0.3);
  padding-bottom: 10px;
  margin-top: 45px;
  margin-bottom: 25px;
}

h3 {
  font-family: "Inter", sans-serif;
  color: #2b2b2b;
  font-weight: 600;
  font-size: 1.1em;
  margin-top: 20px;
}

/* ====== SECTION STYLES ====== */
.exploring, .finding, .insight, .idea {
  border-radius: 12px;
  padding: 18px 24px;
  margin: 20px 0;
  backdrop-filter: blur(10px);
  -webkit-backdrop-filter: blur(10px);
  box-shadow: 0 4px 12px rgba(0,0,0,0.06);
  border: 1px solid rgba(255, 255, 255, 0.4);
}

.exploring {
  background: linear-gradient(135deg, rgba(245, 245, 250, 0.7), rgba(235, 240, 255, 0.6));
  border-left: 4px solid #6366f1;
}

.finding {
  background: linear-gradient(135deg, rgba(240, 249, 255, 0.7), rgba(224, 242, 254, 0.6));
  border-left: 4px solid #0ea5e9;
}

.insight {
  background: linear-gradient(135deg, rgba(220, 252, 231, 0.65), rgba(240, 253, 244, 0.5));
  border-left: 4px solid #1DB954;
  font-weight: 500;
}

.idea {
  background: linear-gradient(135deg, rgba(255, 247, 237, 0.7), rgba(254, 243, 199, 0.6));
  border-left: 4px solid #f59e0b;
  font-weight: 500;
}

.exploring strong, .finding strong, .insight strong, .idea strong {
  display: block;
  font-weight: 700;
  margin-bottom: 8px;
  font-size: 1.05em;
}

/* ====== CODE BLOCKS ====== */
pre {
  background: rgba(250, 250, 250, 0.8);
  border-radius: 8px;
  padding: 16px;
  border: 1px solid rgba(0,0,0,0.05);
  box-shadow: 0 2px 6px rgba(0,0,0,0.04);
  overflow-x: auto;
  color: #1a1a1a;
}

code {
  background: rgba(240, 240, 240, 0.7);
  padding: 2px 6px;
  border-radius: 4px;
  font-size: 0.9em;
  color: #1a1a1a;
}

/* ====== LINKS ====== */
a {
  color: #1DB954;
  text-decoration: none;
  font-weight: 500;
  transition: color 0.2s ease;
}

a:hover {
  color: #1aa34a;
  text-decoration: underline;
}

/* ====== TABLES ====== */
table {
  border-collapse: collapse;
  width: 100%;
  background: rgba(255,255,255,0.7);
  border-radius: 10px;
  overflow: hidden;
  margin: 20px 0;
  box-shadow: 0 2px 8px rgba(0,0,0,0.05);
}

th, td {
  padding: 12px 16px;
  text-align: left;
  border-bottom: 1px solid rgba(0,0,0,0.06);
}

th {
  background: rgba(29, 185, 84, 0.1);
  font-weight: 600;
  color: #004b66;
}

tr:hover {
  background: rgba(29, 185, 84, 0.05);
}

/* ====== TOC FLOATING PANEL ====== */
#TOC {
  font-family: "Inter", sans-serif;
  background: rgba(255, 255, 255, 0.65);
  backdrop-filter: blur(16px) saturate(180%);
  -webkit-backdrop-filter: blur(16px) saturate(180%);
  border-radius: 12px;
  border: 1px solid rgba(255, 255, 255, 0.3);
  box-shadow: 0 8px 24px rgba(0,0,0,0.08);
  padding: 20px;
}

.tocify-header {
  font-weight: 600;
  color: #1DB954;
  font-size: 1.1em;
}

.tocify-item {
  color: #004b66;
  padding: 6px 12px;
  transition: all 0.2s ease;
}

.tocify-item:hover {
  color: #1DB954;
  background: rgba(29, 185, 84, 0.08);
  border-radius: 6px;
}

/* ====== PLOTS ====== */
img {
  border-radius: 10px;
  box-shadow: 0 4px 12px rgba(0,0,0,0.08);
  max-width: 100%;
  height: auto;
}

/* ====== INTRO SECTION ====== */
.intro {
  background: linear-gradient(135deg, rgba(29, 185, 84, 0.12), rgba(30, 215, 96, 0.08));
  border-radius: 12px;
  padding: 25px 30px;
  margin-bottom: 35px;
  border-left: 4px solid #1DB954;
}

.intro p {
  margin: 8px 0;
  font-size: 15px;
}

/* ====== SUMMARY BOX ====== */
.summary-box {
  background: linear-gradient(135deg, rgba(255, 255, 255, 0.8), rgba(240, 248, 255, 0.6));
  border: 2px solid rgba(29, 185, 84, 0.3);
  border-radius: 12px;
  padding: 28px 32px;
  margin: 40px 0;
  box-shadow: 0 4px 12px rgba(0,0,0,0.06);
}

.summary-box h3 {
  margin-top: 0;
  color: #004b66;
  font-size: 1.3em;
  border-bottom: 2px solid rgba(29, 185, 84, 0.2);
  padding-bottom: 8px;
  margin-bottom: 16px;
}

.summary-box ul {
  margin: 12px 0;
  padding-left: 24px;
}

.summary-box li {
  margin: 10px 0;
  line-height: 1.7;
}

.summary-box strong {
  color: #004b66;
  font-weight: 600;
}

/* ====== RESPONSIVE DESIGN ====== */
@media (max-width: 768px) {
  #content {
    margin: 20px auto;
    padding: 30px 25px;
    border-radius: 12px;
  }
  
  h1 {
    font-size: 2em;
  }
  
  h2 {
    font-size: 1.5em;
    margin-top: 35px;
  }
  
  .exploring, .finding, .insight, .idea {
    padding: 14px 18px;
    margin: 16px 0;
  }
  
  .intro, .summary-box {
    padding: 20px 22px;
  }
  
  pre {
    padding: 12px;
    font-size: 0.85em;
  }
  
  table {
    font-size: 0.9em;
  }
  
  th, td {
    padding: 10px 12px;
  }
  
  #TOC {
    padding: 16px;
  }
}

@media (max-width: 480px) {
  #content {
    padding: 20px 18px;
    margin: 10px;
  }
  
  h1 {
    font-size: 1.75em;
  }
  
  h2 {
    font-size: 1.3em;
  }
  
  .exploring, .finding, .insight, .idea {
    padding: 12px 16px;
  }
  
  .intro, .summary-box {
    padding: 18px 20px;
  }
}

/* ====== DARK MODE SUPPORT ====== */
@media (prefers-color-scheme: dark) {
  html, body {
    background: linear-gradient(135deg, #1a1d2e 0%, #16213e 50%, #1a1a2e 100%);
  }
  
  #content {
    color: #e5e7eb;
    background: rgba(30, 30, 45, 0.75);
    border: 1px solid rgba(255, 255, 255, 0.1);
    box-shadow: 0 8px 32px rgba(0, 0, 0, 0.3);
  }
  
  h1 {
    color: #1ed760;
    text-shadow: 0 2px 8px rgba(30, 215, 96, 0.3);
  }
  
  h2 {
    color: #1ed760;
    border-bottom-color: rgba(30, 215, 96, 0.3);
  }
  
  h3 {
    color: #d1d5db;
  }
  
  .exploring {
    background: linear-gradient(135deg, rgba(55, 65, 81, 0.6), rgba(75, 85, 99, 0.5));
    border-left-color: #818cf8;
    color: #e5e7eb;
  }
  
  .finding {
    background: linear-gradient(135deg, rgba(30, 58, 138, 0.5), rgba(37, 99, 235, 0.4));
    border-left-color: #38bdf8;
    color: #e5e7eb;
  }
  
  .insight {
    background: linear-gradient(135deg, rgba(6, 78, 59, 0.5), rgba(4, 120, 87, 0.4));
    border-left-color: #1ed760;
    color: #e5e7eb;
  }
  
  .idea {
    background: linear-gradient(135deg, rgba(120, 53, 15, 0.5), rgba(146, 64, 14, 0.4));
    border-left-color: #fbbf24;
    color: #e5e7eb;
  }
  
  .intro {
    background: linear-gradient(135deg, rgba(6, 78, 59, 0.4), rgba(4, 120, 87, 0.3));
    border-left-color: #1ed760;
    color: #e5e7eb;
  }
  
  .summary-box {
    background: linear-gradient(135deg, rgba(30, 30, 45, 0.8), rgba(37, 99, 235, 0.2));
    border-color: rgba(30, 215, 96, 0.4);
    color: #e5e7eb;
  }
  
  .summary-box h3 {
    color: #1ed760;
    border-bottom-color: rgba(30, 215, 96, 0.3);
  }
  
  .summary-box strong {
    color: #1ed760;
  }
  
  pre {
    background: rgba(240, 245, 250, 0.95);
    border-color: rgba(255, 255, 255, 0.2);
    color: #1a1a1a;
  }
  
  code {
    background: rgba(230, 235, 245, 0.9);
    color: #1a1a1a;
  }
  
  pre code {
    background: transparent;
    color: #1a1a1a;
  }
  
  table {
    background: rgba(30, 30, 45, 0.7);
    border-color: rgba(255, 255, 255, 0.1);
    color: #e5e7eb;
  }
  
  th {
    background: rgba(6, 78, 59, 0.3);
    color: #1ed760;
  }
  
  td {
    border-bottom-color: rgba(255, 255, 255, 0.05);
  }
  
  tr:hover {
    background: rgba(30, 215, 96, 0.08);
  }
  
  a {
    color: #1ed760;
  }
  
  a:hover {
    color: #1aa34a;
  }
  
  #TOC {
    background: rgba(30, 30, 45, 0.75);
    border-color: rgba(255, 255, 255, 0.1);
    box-shadow: 0 8px 24px rgba(0, 0, 0, 0.3);
  }
  
  .tocify-header {
    color: #1ed760;
  }
  
  .tocify-item {
    color: #d1d5db;
  }
  
  .tocify-item:hover {
    color: #1ed760;
    background: rgba(30, 215, 96, 0.1);
  }
  
  img {
    border-color: rgba(255, 255, 255, 0.1);
    box-shadow: 0 4px 12px rgba(0, 0, 0, 0.3);
  }
}
```

<div id="content">

# Spotify 2023 Streaming Analysis

<div class="intro">
**Statistical analysis of the most streamed songs on Spotify in 2023**

Exploring patterns in music characteristics, streaming success, and temporal trends using R statistical methods. All insights are observational and do not imply causation.

**Dataset:** [Most Streamed Spotify Songs 2023](https://www.kaggle.com/datasets/nelgiriyewithana/top-spotify-songs-2023/data) from [Kaggle](https://www.kaggle.com/)  
**Tools:** R, ggplot2, dplyr, e1071
</div>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6,
  fig.align = 'center'
)
```

```{r load-packages, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(e1071)
load("spotify-2023-clean.Rdata")
```

---

## Executive Summary

<div class="summary-box">

### Key Findings at a Glance

**Temporal Trend**  
Songs released from 2020 onward show significantly higher danceability (p < 0.001), indicating a measurable shift in music production toward social listening contexts.

**Tempo & Popularity**  
Slower-tempo tracks (BPM < 100) achieve highest median streams, suggesting broad appeal across diverse listening scenarios including focus, relaxation, and background contexts.

**Feature Relationships**  
Energy and valence demonstrate moderate positive correlation (r ≈ 0.43). Danceability and valence show strongest correlation among analyzed features. BPM operates independently of mood-related metrics.

**Mode Effects**  
Major and minor modes create statistically robust differences in energy, danceability, and valence, surviving multiple testing corrections.

**Top Content Profile**  
The highest-streaming songs maintain moderate-to-high positivity (valence > 43%), establishing a baseline characteristic for commercially successful content.

### Strategic Recommendations

* Prioritize recent high-danceability content in discovery algorithms
* Feature slower-tempo tracks in context-based playlists (study, focus, chill)
* Leverage mode distinctions for emotional playlist curation
* Maintain valence >40% threshold for broad-audience editorial selections
* Keep BPM as independent preference filter due to weak correlation with mood features

</div>

---

## Detailed Analysis

## Danceability Distribution

<div class="exploring">
**Research Questions:**

* What is the mean danceability level?
* Which statistical distribution best describes danceability levels?
</div>

```{r danceability-analysis}
summary(data)
danceability_mean <- summary(data$danceability)[c("Mean")]
danceability_median <- summary(data$danceability)[c("Median")]
danceability_std <- sd(data$danceability)

hist(data$danceability,
     main = "Distribution of Danceability",
     xlab = "Danceability (%)",
     ylab = "Frequency",
     col = "#1DB954",
     border = "white",
     breaks = 20)

cat("Mean:", round(danceability_mean, 2), "\n")
cat("Median:", round(danceability_median, 2), "\n")
cat("Std Dev:", round(danceability_std, 2), "\n")
cat("Skewness:", round(skewness(data$danceability), 2), "\n")
```

<div class="finding">
**Findings:**

* Mean danceability: *66.98%*
* Distribution: *Left-skewed* (most songs have moderate-to-high danceability)
</div>

<div class="insight">
**Insight:**

The majority of 2023's top songs have medium-high danceability scores, suggesting Spotify is increasingly used in social and party contexts where danceable music is preferred.
</div>

<div class="idea">
**Product Implication:**

Curating and promoting more party-themed and workout playlists could align with user behavior and increase session length.
</div>

---

## Energy vs. Valence Relationship

<div class="exploring">
**Research Questions:**

* What is the relationship between energy and valence?
* Do release years show any trend in this relationship?
</div>

```{r energy-valence}
ggplot(data, aes(x = energy, y = valence, color = released_year)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = lm, se = FALSE, color = "black", linewidth = 1.2) +
  scale_color_gradient(low = "#FF6347", high = "#1DB954", name = "Release Year") +
  labs(
    title = "Energy vs. Valence Across Release Years",
    x = "Energy (%)",
    y = "Valence (Positivity %)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    legend.position = "right"
  )

cor_energy <- cor.test(data$energy, data$valence)
cat("Correlation between energy and valence:", round(cor_energy$estimate, 3), "\n")
cat("P-value:", format(cor_energy$p.value, scientific = TRUE), "\n")
```

<div class="finding">
**Findings:**

* *Positive correlation* between energy and valence (r ≈ 0.43)
* No clear temporal trend across release years
</div>

<div class="insight">
**Insight:**

Energetic songs tend to be more positive, but the relationship is moderate. This suggests energy and valence capture different aspects of musical mood.
</div>

<div class="idea">
**Product Implication:**

Keep energy and valence as separate metrics in recommendation algorithms to capture nuanced mood preferences and improve playlist personalization.
</div>

---

## BPM Impact on Streaming Success

<div class="exploring">
**Research Questions:**

* Which BPM (Beats Per Minutes) category shows highest streaming numbers?
* How does tempo relate to popularity?
</div>

```{r bpm-analysis}
data <- mutate(data,
  BPM_category = case_when(
    bpm < 100 ~ "Slow",
    bpm <= 120 ~ "Medium",
    bpm > 120 ~ "Fast"
  )
)

data$BPM_category <- factor(data$BPM_category, levels = c("Slow", "Medium", "Fast"))

ggplot(data, aes(x = BPM_category, y = streams, fill = BPM_category)) +
  geom_boxplot(alpha = 0.8, outlier.color = "#FF6347", outlier.size = 2) +
  geom_hline(yintercept = median(data$streams[data$BPM_category=="Slow"]), linetype = "dotted") +
  scale_fill_manual(
    values = c("Slow" = "#1ed760", "Medium" = "#1ed760", "Fast" = "#1ed760")
  ) +
  scale_y_log10(labels = scales::comma) +
  labs(
    title = "Streaming Distribution by BPM Category",
    x = "BPM Category",
    y = "Number of Streams (log scale)"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 16),
    legend.position = "none"
  )
```

<div class="finding">
**Findings:**

* *Slow BPM* (<100) has highest median streams
* Lower tempo consistently associated with higher popularity
</div>

<div class="insight">
**Insight:**

Slower-paced tracks dominate streaming success, possibly because they work well for diverse contexts (background listening, focus, relaxation).
</div>

<div class="idea">
**Product Implication:**

Feature slower-tempo tracks prominently in broad-audience playlists and context-based recommendations (study, chill, focus).
</div>

---

## Audio Feature Correlation Analysis

<div class="exploring">
**Research Questions:**

* Which audio features are most strongly correlated?
* Which features show independence?
</div>

```{r correlation-analysis}
correlation_matrix <- cor(
  data[, c("danceability", "energy", "valence", "bpm")],
  use = "complete.obs"
)

heatmap(correlation_matrix,
  symm = TRUE,
  main = "Audio Feature Correlation Heatmap",
  col = colorRampPalette(c("white", "#1DB954"))(20),
  margins = c(10, 10)
)
```

<div class="finding">
**Findings:**

* **Strongest correlation:** Danceability ↔ Valence
* **Weakest correlation:** BPM with other features
</div>

<div class="insight">
**Insight:**

Danceable songs tend to be more positive. BPM operates independently, meaning tempo preferences vary significantly across listeners regardless of other mood factors.
</div>

<div class="idea">
**Product Implication:**

Use danceability-valence correlation for mood-based playlist generation. Keep BPM as an independent filter for user preferences.
</div>

---

## Temporal Trends: Danceability Over Time

<div class="exploring">
**Research Questions:**

* Has danceability changed significantly over time?
* Are songs from 2020+ more danceable?
</div>

```{r temporal-analysis}
data <- mutate(data,
  release_period = case_when(
    released_year < 2020 ~ "Before 2020",
    released_year >= 2020 ~ "2020 or later"
  )
)

t_test_result <- t.test(danceability ~ release_period, data = data)

boxplot(danceability ~ release_period,
  data = data,
  main = "Danceability: Before vs. After 2020",
  xlab = "Release Period",
  ylab = "Danceability (%)",
  col = c("#FF6347", "#1DB954"),
  border = "black",
  horizontal = FALSE
)

cat("T-test p-value:", format(t_test_result$p.value, scientific = TRUE), "\n")
cat("Mean (Before 2020):", round(t_test_result$estimate[1], 2), "\n")
cat("Mean (2020+):", round(t_test_result$estimate[2], 2), "\n")
```

<div class="finding">
**Findings:**

* P-value: *6.15e-12* (highly significant)
* Songs from 2020+ have *significantly higher danceability*
</div>

<div class="insight">
**Insight:**

There's a clear shift toward more danceable music in recent years, reflecting evolving listener preferences for energetic, social listening experiences.
</div>

<div class="idea">
**Product Implication:**

Update algorithmic weights to favor recent high-danceability tracks in discovery playlists to match current user preferences.
</div>

---

## Major vs. Minor Mode Comparison

<div class="exploring">
**Research Questions:**

* Do major and minor mode songs differ across multiple audio features?
* Which differences remain significant after multiple testing correction?
</div>

```{r mode-analysis}
variables <- c("energy", "danceability", "valence", "acousticness", 
               "speechiness", "liveness", "bpm", "instrumentalness")

my_t_test <- function(variable) {
  t.test(data[[variable]] ~ data$mode)
}

t_test_results <- lapply(variables, my_t_test)
p_values <- sapply(t_test_results, function(test) test$p.value)

results <- data.frame(
  Variable = variables,
  P_Value = round(p_values, 5),
  Bonferroni = round(p.adjust(p_values, method = "bonferroni"), 5),
  BH = round(p.adjust(p_values, method = "BH"), 5)
)

knitr::kable(results, caption = "Multiple Testing Results: Major vs. Minor Mode")

cat("\nSignificant tests (α = 0.05):\n")
cat("Before correction:", sum(results$P_Value < 0.05), "\n")
cat("After Bonferroni:", sum(results$Bonferroni < 0.05), "\n")
cat("After BH:", sum(results$BH < 0.05), "\n")
```

<div class="finding">
**Findings:**

* *3 features* significant before correction
* *2 features* remain significant after Bonferroni correction
* Major mode songs show higher energy, danceability, and valence
</div>

<div class="insight">
**Insight:**

Musical mode creates robust differences in perceived mood and energy, even after strict statistical corrections.
</div>

<div class="idea">
**Product Implication:**

Use mode as a distinguishing factor: Major mode for upbeat/party playlists, Minor mode for emotional/calm playlists.
</div>

---

## Top 100 Songs: Valence Analysis

<div class="exploring">
**Research Question:**

* What is the typical positivity level of the most popular songs?
</div>

```{r bootstrap-analysis}
top_100 <- data %>%
  arrange(desc(streams)) %>%
  slice(1:100)

bootstrap_sample_mean <- function(data) {
  bootstrap_sample <- sample(data, length(data), replace = TRUE)
  return(mean(bootstrap_sample))
}

set.seed(123)
bootstrap_means <- replicate(1000, bootstrap_sample_mean(top_100$valence))
ci <- quantile(bootstrap_means, c(0.025, 0.975))

hist(bootstrap_means,
  main = "Bootstrap Distribution of Mean Valence (Top 100 Songs)",
  xlab = "Mean Valence (%)",
  col = "#1DB954",
  border = "white",
  breaks = 30
)
abline(v = ci, col = "#FF6347", lwd = 2, lty = 2)

cat("95% Bootstrap Confidence Interval:\n")
cat("Lower bound:", round(ci[1], 2), "%\n")
cat("Upper bound:", round(ci[2], 2), "%\n")
```

<div class="finding">
**Findings:**

* 95% CI lower bound: *43.69%*
* Top songs maintain moderate-to-high positivity
</div>

<div class="insight">
**Insight:**

The most streamed songs cluster around moderate-to-high valence, suggesting mainstream success favors emotionally positive content.
</div>

<div class="idea">
**Product Implication:**

Prioritize tracks with valence >40% in editorial playlists targeting broad audiences to maximize engagement.
</div>

---

## Conclusions & Next Steps

<div class="summary-box">

### Strategic Insights Recap

This analysis reveals five actionable patterns in 2023's top-streaming music:

1. **Danceability is trending upward** — Modern releases show significantly higher scores
2. **Slower tempo drives engagement** — BPM <100 correlates with higher streams
3. **Energy-valence correlation exists but is moderate** — Both metrics provide value
4. **Major mode signals upbeat content** — Statistically significant after corrections
5. **Popular content clusters around positive valence** — Above 43% threshold

### Recommended Next Steps

**For Product Development:**

* A/B test party-focused playlist promotions based on high-danceability filtering
* Implement tempo-based segmentation in personalization algorithms
* Build temporal trend dashboards for content strategy teams
* Develop mode-aware playlist generation features

**For Further Analysis:**

* Investigate regional differences in danceability preferences
* Analyze seasonal patterns in audio feature trends
* Examine artist-level consistency across these metrics
* Study playlist context impact on feature importance

</div>

</div>